A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2012-05-12 papers/contact/paper.tex
Fix bug in ContactDensity (the "Mysterious Four") and a few minor paper changes Tests skipped!
clean up appendices of contact paper a bit Tested everything on bingley.

\fixme{Analogous to the assymetric case is the symmetric, for which 
we are no longer able to simply pick a hypothetical sphere centered 
at position $r$ and expand its radius.  Instead we must formulate a way 
to expand the radii of spheres that just touch a specific point of 
contact and only expand them at that point of contact, in a symmetric 
fashion.  Therefore, when we compute the neccasary derivitives wih 
respect to $R$', we expand a multitude of spheres, all of them positioned 
around and just touching the point $\mathbf{r}$.  When their radii 
expand, each does not expand outward in all directions as in the 
case of A, rather they each expand outward at only the point $\mathbf{r}$.  One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains 
at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
in this peculiar fashion, we have in the end the same amount of shell 
area protruding outward as we would if we had simply expanded one 
sphere in fashion of case A.  Also the protrusions, once added up 
for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
in their direction of protrusion.  Each protruding sphere bit will 
expand outward from its center and all directions will be covered uniformly.

@@

In the case of A, the changing radius we refer to is the radius of a sphere that is centered at $\mathbf{r}$.  The $n_3$ values that will be affected are those which are a distance $R$ from $\mathbf{r}$.

\begin{align}
  n_3^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}'')) d\mathbf{r}''\\ 
  \frac{\delta n_3^{A} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}'' \\
   &= n (\mathbf{r}) \delta(|\mathbf{r'} - \mathbf{r}| - R(\mathbf{r}))
\end{align}

And likewise:
\begin{align}
  n_2^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) d \mathbf{r}''\\
  \frac{\delta n_2^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))\\
 \end{align}

where the trouble here is that the derivative of the delta function,
$\delta'$, isn't something we really like to deal with.  Particularly
as we're looking  at a 3D integral involving a 1D delta function,
which means there are Jacobian factors to be considered.  On the other
hand, there's no reason why we couldn't convolve with this kernel as
efficiently as we do with any other kernel.
\\
\\
The derivatives of the remaining scalar densities can be reduced to
sums of the terms above:
\begin{align}
  n_1(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})}
  &= -\frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
  -
  \frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
\end{align}
\begin{align}
  n_0(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})^2}
  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})^2}
  &= -\frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})^2}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
  -
  2\frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})^3}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
\end{align}

Since the vector-weighted densities have no additional occurrences of
$R$, their derivatives come out similarly.
\begin{align}
  \mathbf{n}_{V2}^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) 
    \frac{\mathbf{r'}-\mathbf{r''}}{|\mathbf{r'}-\mathbf{r''}|} d \mathbf{r}''\\
  \frac{\delta \mathbf{n}_{V2}^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= - n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))
    \frac{\mathbf{r}'-\mathbf{r}}{|\mathbf{r}'-\mathbf{r}|}\\
\end{align}

and analogously for $\mathbf{n}_{V1}$:

\begin{align}
  \mathbf{n}_{V1}(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}\\
  \frac{\delta \mathbf{n}_{V1}(\mathbf{r}')}{\delta R(\mathbf{r})}
  &= -\frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}
  -
  \frac{n(\mathbf{r})}{4\pi
    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}  
\end{align}

So, what is this convolution with a $\delta'$? Let's look at the
scalar version:
\fixme{With following the primes are certainly not in the right places}
\begin{align}
  \bar{f}(\mathbf{r})
  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
  \\
  &= \int
        f(\mathbf{r'})
        \lim_{\Delta R \rightarrow 0}
          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
            -
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &= \lim_{\Delta R \rightarrow 0}
          \frac{\int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
            -
            \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &\approx\lim_{\Delta R \rightarrow 0}
          \frac{\int
            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
            -
            \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &\approx\lim_{\Delta R \rightarrow 0}
          \frac{2\frac{\Delta R}{R} \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &= \frac{2}{R} \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
\end{align}
There is a sign ambiguity here with regard to whether we are asking
about the derivative of this delta function with regard to $R$ or
something else.  We just need to get it right.

@@

For the derivation of $n_{3}$ in the symmetric case:

\begin{align}  
  n_3^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}')) d\mathbf{r}''\\
  \frac{\delta n_3^{S} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}''
\end{align}

\fixme{Not sure if we need any more explanation here (probably want better than this) -  $\delta R(\mathbf{r})$ in this equation refers to an expansion, in the fashion described above, of all the spheres that just touch $\mathbf{r}$.  We can see that if the radii change in this fashion, the only point at which $n_3(\mathbf{r}')$ will be affected is at $\mathbf{r}' = \mathbf{r}$, so that $\delta n_3(\mathbf{r}')$ is zero everywhere except where $\mathbf{r}' = \mathbf{r}$.  Thus we have the 2nd $\delta$ function in the lower integral.}
\\
\\
This term in the free energy functional, once integrated, takes an interesting form:

\begin{align}
  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} 
  \frac{\delta n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}' 
  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}'' d\mathbf{r}' \\
  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r} - \mathbf{r}''| - R(\mathbf{r})) d\mathbf{r}''\\
  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}'')} n_2(\mathbf{r})\\
\end{align}

 $n_2(\mathbf{r}')$ in the case of S in calculated similarly
\begin{align}
  n_2^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
    \frac{\delta n_2^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int n(\mathbf{r}'') 
  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}''
\end{align}

And similarly to the A case we compute
\begin{align}
  n_1^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
    \frac{\delta n_1^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
  -
  \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
\end{align}

and 
\begin{align}
  n_0^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
    \frac{\delta n_0^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} 
  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
  -
  2\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^3} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
\end{align}

\begin{align}
  \mathbf{n}_{V2}^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'))
  \frac{(\mathbf{r}' - \mathbf{r}'')}{|\mathbf{r}' - \mathbf{r}''|} d \mathbf{r}''\\
  \frac{\delta \mathbf{n}_{V2}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= \delta(\mathbf{r} - \mathbf{r}')
  \int n(\mathbf{r}'') \delta'(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) 
  \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} d\mathbf{r}''
\end{align}

\begin{align}
 \mathbf{n}_{V1}^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}\\
 \frac{\delta \mathbf{n}_{V1}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
   \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} 
   -
   \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r'' 
   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}
\end{align}

\begin{widetext}
\section{Formulas generated from Haskell}

@@

\fixme{This is the hypthetical end of the paper}

 \fixme {
And then there's a lot with taking the derivative of the delta for the case of $n_{2}^{S}(\mathbf{r}')$.  Not sure if we'll use any of this, but wrote down some equations that Roundy wrote out.  So can use the terms at least if going to write about this - I'm just copying it in for now (probably a lotof mistakes)}

\begin{align}
  p(\mathbf{r}) &= \frac{\delta A}{\delta n_2(\mathbf{r}')} \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}'\\
  &= \frac{\delta A}{\delta n_2(\mathbf{r})} - \int n(\mathbf{r}'') \delta' (|\mathbf{r} - \mathbf{r}''| - R) d\mathbf{r}''
  \int n (\mathbf{r} +\mathbf{r}'') \delta'(|\mathbf{r}'''| - R) d \mathbf{r}'''\\
  &= \int n(\mathbf{r} + \mathbf{r}''') \delta'(|\mathbf{r}'''| - R) r'''^2 cos\theta d \theta''' dr'''\\
  du = \delta'(\mathbf{r}''' - R) d \mathbf{r}'''     v = n(\mathbf{r} - \mathbf{r}'')r'''\\
   u = \delta(\mathbf{r}''' - R)                       
   dv = \mathbf{\nabla} n(\mathbf{r} + \mathbf{r}'') \cdot \frac{\mathbf{r}'''}{r'''} + 2n(\mathbf{r}' - \mathbf{r}'')r'''\\
  &= -\int(\mathbf{\nabla}n(\mathbf{r}' + \mathbf{r}'') \cdot \mathbf r'''(hat)r'''^2 dr''' 
  + 2n(\mathbf{r} + \mathbf{r}''')r''') dr''' \delta(\mathbf{r} - R) dcos\theta''' d\theta''\\
  &= - \int \delta(\mathbf{r}''' - R) d\mathbf{r}''' 
  (\mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) + \frac{2n(\mathbf{r} + \mathbf{r}'')}{\mathbf{r}'''})\\
  &= \frac{2n_2(\mathbf{r})}{R} + 
  \int \delta(|\mathbf{r}'''| - R) \mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) d \mathbf{r}'''\\
\end{align}

\end{widetext}
\end{widetext}

>>>>

\fixme{Analogous to the assymetric case is the symmetric, for which 
we are no longer able to simply pick a hypothetical sphere centered 
at position $r$ and expand its radius.  Instead we must formulate a way 
to expand the radii of spheres that just touch a specific point of 
contact and only expand them at that point of contact, in a symmetric 
fashion.  Therefore, when we compute the neccasary derivitives wih 
respect to $R$', we expand a multitude of spheres, all of them positioned 
around and just touching the point $\mathbf{r}$.  When their radii 
expand, each does not expand outward in all directions as in the 
case of A, rather they each expand outward at only the point $\mathbf{r}$.  One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains 
at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
in this peculiar fashion, we have in the end the same amount of shell 
area protruding outward as we would if we had simply expanded one 
sphere in fashion of case A.  Also the protrusions, once added up 
for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
in their direction of protrusion.  Each protruding sphere bit will 
expand outward from its center and all directions will be covered uniformly.
See Figure \ref{fig:contact}.

\begin{figure}
\includegraphics[width=5cm]{figs/contact}
\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
  consist of those which just touch the point $\mathbf{r}$.}
\label{fig:contact}
\end{figure}

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^A(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position of the sphere in
question---which is to say, the location where the density itself is
evaluated.  This is the most natural way to construct such a
functional, in which the radius of spheres is a function of position.
Thus we define $n_3^A$ as
\begin{align}
  n_3^{A}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr')) d\rr'
\end{align}
and we define the other weighted densities similarly.  In order to
compute derivatives of the free energy with respect to $R(\rr)$, we
need to find $g_\sigma^A$.  In the case of $n_3^A$, we needS
\begin{align}
  \frac{\delta n_3^{A} (\rr')}{\delta R(\rr)} &=
  \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'')) \delta(\rr-\rr'') d\rr'' \\
   &= n (\rr) \delta(|\rr' - \rr| - R(\rr))
\end{align}
Similarly, for $n_2^A$, we find
\begin{align}
  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - R(\rr'')) d \rr''\\
  \frac{\delta n_2^{A}(\rr')}{\delta R(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - R(\rr))\\
\end{align}
This is somewhat more challenging, since it involves an unfamiliar
function, the derivative the Dirac delta function $\delta'$.
%
The derivatives of the remaining scalar densities can be reduced to
sums of the terms above:
\begin{align}
  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')}
  \delta(|\rr'-\rr''| - R(\rr'')) \\
  \frac{\delta n_1(\rr')}{\delta R(\rr)}
  &= -\frac{n(\rr)}{4\pi
    R(\rr)}\delta'(|\rr'-\rr| - R(\rr))
  -
  \frac{n(\rr)}{4\pi
    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr))
\end{align}
and
\begin{align}
  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')^2}
  \delta(|\rr'-\rr''| - R(\rr'')) \\
  \frac{\delta n_1(\rr')}{\delta R(\rr)^2}
  &= -\frac{n(\rr)}{4\pi
    R(\rr)^2}\delta'(|\rr'-\rr| - R(\rr))
  -
  2\frac{n(\rr)}{4\pi
    R(\rr)^3}\delta(|\rr'-\rr| - R(\rr))
\end{align}
The vector-weighted densities have no additional occurrences of $R$,
their functional derivatives come out similarly:
\begin{align}
  \mathbf{n}_{V2}^{A}(\rr) &= \int n(\rr') \delta(|\rr' - \rr''| - R(\rr''))
    \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta R(\rr)} &= - n(\rr) \delta'(|\rr' - \rr| - R(\rr))
    \frac{\rr'-\rr}{|\rr'-\rr|}\\
\end{align}
and analogously for $\mathbf{n}_{V1}^A$:
\begin{align}
  \mathbf{n}_{V1}^A(\rr) &= \int d\rr' \frac{n(\rr')}{4\pi R(\rr')}
  \delta(|\mathbf{r}-\rr'| - R(\rr')) \frac{\rr-\rr'}{|\rr-\rr'|}\\
  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta R(\rr)}
  &= -\frac{n(\rr)}{4\pi
    R(\rr)}\delta'(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
  -
  \frac{n(\rr)}{4\pi
    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
\end{align}
\fixme{double-check the directions of all the vector quantities...}

%% So, what is this convolution with a $\delta'$? Let's look at the
%% scalar version:
%% \fixme{With following the primes are certainly not in the right places}
%% \begin{align}
%%   \bar{f}(\mathbf{r})
%%   &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
%%   \\
%%   &= \int
%%         f(\mathbf{r'})
%%         \lim_{\Delta R \rightarrow 0}
%%           \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
%%             -
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%%       }{\Delta R}
%%   \\
%%   &= \lim_{\Delta R \rightarrow 0}
%%           \frac{\int
%%             f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
%%             -
%%             \int
%%             f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%%       }{\Delta R}
%%   \\
%%   &\approx\lim_{\Delta R \rightarrow 0}
%%           \frac{\int
%%             \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%%             -
%%             \int
%%             f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%%       }{\Delta R}
%%   \\
%%   &\approx\lim_{\Delta R \rightarrow 0}
%%           \frac{2\frac{\Delta R}{R} \int
%%             f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%%       }{\Delta R}
%%   \\
%%   &= \frac{2}{R} \int
%%             f(\mathbf{r'})
%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
%% \end{align}
%% There is a sign ambiguity here with regard to whether we are asking
%% about the derivative of this delta function with regard to $R$ or
%% something else.  We just need to get it right.

This completes our exploration of the partial derivatives of weighted
densities of the $A$ variety.

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^S(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position at which the weighted
density is being evaluated.  This is less intuitive than the $A$ case,
since \emph{nothing is there}!  However, in practice it corresponds to
the picture shown in Figure~\ref{fig:contact}, which involves
extending the radius for spheres just at the position touching a
given spot, where we are evaluating the correlation function.  We
define the weighted density $n_3^S$ as:
\begin{align}
  n_3^{S}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr)) d\rr'
\end{align}
In the symmetric case, the functional derivative is strikingly simple:
\begin{align}
  \frac{\delta n_3^{S} (\rr')}{\delta R(\rr)} &=
   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'))
   \delta(\rr'-\rr) d\rr''
   \\
   &= \delta(\rr'-\rr) n_2^S(\rr')
\end{align}
The functional derivative does \emph{not} eliminate the integral,
which simply changes from one weighted density to another.  The
corresponding term in the free energy functional, once integrated,
takes a simple form:
\begin{align}
  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
  \frac{\delta n_3^{S}(\rr')}{\delta R(\rr)} d\rr'
  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
  n_2^S(\rr') d\rr' \\
  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
\end{align}
which means that we can compute $g_\sigma^S(\rr)$ using only
convolutions of the density, in strong contrast with
$g_\sigma^A(\rr)$, which requires convolutions of derivatives of the
free energy density.

The functional derivative of $n_2^S(\rr')$ is quite similar:
\begin{align}
  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| + R(\rr))d\mathbf
  r'\\
  \frac{\delta n_2^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
  \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
\end{align}
where we recognize that the convolution---while not one of the standard
FMT weighted density---is a simple convolution of the density.  We can
go on to compute the derivatives of the remaining weighted densities:
\begin{align}
  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
    \frac{\delta n_1^{S}(\rr')}{\delta R(\rr)} &=
    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)}
    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
    + n_0^s(\rr') \right)
\end{align}
\begin{align}
  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
    \frac{\delta n_0^{S}(\rr')}{\delta R(\rr)}
    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)^2}
    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
    +
    \frac{2}{R} n_0^S(\rr') \right)
\end{align}
\begin{align}
  \mathbf{n}_{V2}^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))
  \frac{(\rr - \rr')}{|\rr - \rr'|} d \rr'\\
  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= \delta(\rr - \rr')
  \int n(\rr'') \delta'(|\rr' - \rr''| - R(\rr'))
  \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
\end{align}
\begin{align}
 \mathbf{n}_{V1}^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'
   \frac{(\rr - \rr')}{|\rr - \rr'|}\\
 \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta R(\rr)}
 &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi R(\rr)}
   \delta'(|\rr'-\rr''| - R(\rr')) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
   + \mathbf{n}_{V2}^S(\rr') \right)
\end{align}

The above equations involve just two additional weighted densities
beyond those required for FMT: one scalar density computed with a
derivative of the Dirac delta function, and one analogous vector
density.

\section{Formulas generated from Haskell}

@@

\end{widetext}
